Assignment 1 
Pablo Tercero
========================================================

This is the R Markdown document for Assignment 1 on Reproducible Research's coursera course.I present below the steps taken to produce the answer to the assignment questions:

**Loading the data**
First I loaded the data and looked at the nature of the data

```{r}
actDat <- read.csv("activity.csv",header = TRUE)
head(actDat)
str(actDat)
```

```
Then 1st question:
**What is mean total number of steps taken per day?**

First I aggregate the data by date getting the total steps by date and then I plot them 
To get the mean and median I use summary
```{r}
stepsByDay <- with(actDat,aggregate(steps,by=list(date),FUN="sum"))
hist(stepsByDay$x)
summary(stepsByDay$x)

```

**What is the average daily activity pattern?**
I use tapply to get the mean steps by interval and then I plot using line (type='l')
The max is obtained with which.max

```{r}
stepsByInterval <- tapply(actDat$steps,actDat$interval,mean,na.rm=TRUE)
plot(stepsByInterval,type = 'l')
which.max(stepsByInterval) #it's interval 104
```

**Imputing missing values**
This took a while but doing some reseach I found an R-bloggers post about this great process/tool for imputing missing data
http://www.r-bloggers.com/imputing-missing-data-with-r-mice-package/

I loaded the necessary library and followed steps to get the # of NAs 
These can also be obtained with 
```{r}
summary(actDat)
```
that gave us 2304 NAs, but the package was really useful to find a method to impute values for NAs
```{r}
library(mice)
md.pattern(actDat)
```

The method is mice , it uses method pmm (predictive mean matching) but there are other methods
There's a nice paper at : http://www.stefvanbuuren.nl/publications/MICE%20in%20R%20-%20Draft.pdf
(the code took a long time to evaluate so I only use 2 iterations in this file but I used 50 to run my own code , tests , the more iterations the better)
```{r}
tempData <- mice(actDat,maxit = 2,method = 'pmm',seed = 500)
summary(tempData)
```
tempData$imp$steps shows the steps and results for each iteration, I don't ouput that either

```{r}
completeData <- complete(tempData,1)
head(completeData)
```

completeData now has all of the data without the NAs and is the new data set to compare with the set with NAs

Using the exact same code with a different data set , I get the plot and the mean,median

```{r}
stepsByDayComp <- with(completeData,aggregate(steps,by=list(date),FUN="sum"))
hist(stepsByDayComp$x)
summary(stepsByDayComp$x)
```

**Are there differences in activity patterns between weekdays and weekends?**

 I noticed that the date field needs to be converted to data , so I use lubridate to convert
```{r}
library(lubridate)
actDat$date <- ymd(actDat$date)
```

The steps below accomplish the following 
1. I get a new data set to manipulate the data , create weekend info etc
2. I create two new columns 
  + WeekD is the weekday based on the date info 
  + Weekend which is a Y for weekend, N for "weekday"" based on what weekday(the variable) is
3. I find the mean steps for each interval  for weekend and weekdays
4. I rename the columns of the data frame (obtained with aggregate , mean) with better names

I do head and tail of the data sets created to check if the variables were added correctly

```{r}
complDat <- completeData
complDat$WeekD <- weekdays(ymd(complDat$date))
complDat$Weekend = ifelse(complDat$WeekD %in% c("Saturday","Sunday"),"Y","N")
head(complDat)
tail(complDat)
complstepsByIntervalAgg <- aggregate(complDat$steps,list(complDat$interval,as.factor(complDat$Weekend)),mean,na.rm=TRUE)
colnames(complstepsByIntervalAgg) <- c('interval','Weekend','stepsMean')
head(complstepsByIntervalAgg)
tail(complstepsByIntervalAgg)
```

Finally I make a ggplot (line plot) with facets to represent the difference in mean steps on weekdays and weekends
```{r}
library(ggplot2)
ggplot(complstepsByIntervalAgg,aes(interval,stepsMean))+geom_line()+facet_grid(.~Weekend)
```

This completes the assignment
